---
title: "Business Analytics Assignment-1"
author: "Spandana Sodadasi"
date: "2023-09-17"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment = NULL)
```

#Installed the ISLR library using the install.packages() command and loading the package by using the library function.
```{r}
library(ISLR)
library(ggplot2)
View(Carseats)
```

#Printing the summary of the Carseats dataset.
```{r}
print(summary(Carseats))
```

#Understanding the data structure of carseats.
```{r}
names(Carseats)
```

```{r}
str(Carseats)
```

```{r}
head(Carseats)
```

#Number of Observation(rows) present in the dataset.
```{r}
nrow(Carseats)
```

#Maximum value of the advertising attribute.
```{r}
max(Carseats$Advertising)
```

#Calculating the IQR of the Price attribute.
```{r}
print(paste("Price IQR is", IQR(Carseats$Price)))
```

#Plotting the Sales variable aganist the Price variable.
```{r}
ggplot(Carseats, aes(x = Price, y = Sales), main = "Scatter Plot") +
  geom_point() +
  labs(title = "Scatter Plot of Sales vs Price",
       x = "Price", y = "Sales") +
  theme_minimal() +
  geom_smooth(method = "lm", formula = y ~ x, color = "red",linetype = "dashed", se=FALSE)+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

```

In the scatter plot we can see a mild negative relationship between Price and Sales variables. In the further step we will calculate the correlation coefficient to have a better understanding of this relationship.

#Calculating the correlation between Price and Sales attributes.
```{r}
cor(Carseats$Price,Carseats$Sales)
```

A correlation coefficient is -0.4449507 and the sign indicates a negative correlation between Price and Sales attributes. This means that if Price increases, Sales will tend to decrease.

